Assessing the Accessibility of Protein Tunnels using Motion Planning

### 2017-12-08 paper.tex
intro correction

Reactivity of proteins with other molecules plays a crucial role in many research disciplines, including protein engineering and drug design.
Protein reactivity is tightly connected with the presence of the void space between its amino acids. 
This space, forming a path called tunnel, can be utilized by a small molecule (ligand) which can enter the protein inner space and react with its amino acids surrounding a specific site, called the active site~\cite{gora2013gates}.
The size and location of tunnels determine the ligand size and shape which can enter the active site.
Currently the biochemists are using various computational methods for tunnel detection (Fig.~\ref{fig::motiv}) in order to estimate the possibility of interactions between a protein and a ligand.
This helps them to reduce the number of necessary in-vitro lab experiments.
The most common methods detect the tunnels using Voronoi diagrams~\cite{yaffe2008,caver3}, assuming a spherical ligand (probe), i.e., the approximation of the ligand by a bounding sphere.
\textcolor{blue}{However, the ligands are typically of a non-spherical shape which is completely ignored by these methods. 
Additionally, when studying the tunnel evolution and behavior over time, the traditional methods detect tunnels in individual time frames and perform their clustering to reveal their temporal correspondence. 
Therefore, they do not take into account the ligand position in the previous step, which leads to situations, when the information about a given tunnel is completely missing in few or many frames.}

@@

Our proposed method was tested by the biochemists who compared the results with trajectories of ligands determined experimentally in the lab.
\textcolor{blue}{In the following we first mention several approaches and methods related to our solution and then describe our proposed method in detail.}

>>>>

Reactivity of proteins with other molecules plays a crucial role in many research disciplines, including protein engineering and drug design.
Protein reactivity is tightly connected with the presence of the void space between its amino acids. 
This space, forming a path called tunnel, can be utilized by a small molecule (ligand) which can enter the protein inner space and react with its amino acids surrounding a specific site, called the active site~\cite{gora2013gates}.
The size and location of tunnels determine the ligand size and shape which can enter the active site.
Currently the biochemists are using various computational methods for tunnel detection (Fig.~\ref{fig::motiv}) in order to estimate the possibility of interactions between a protein and a ligand.
This helps them to reduce the number of necessary in-vitro lab experiments.
The most common methods detect the tunnels using Voronoi diagrams~\cite{yaffe2008,caver3}, assuming a spherical ligand (probe), i.e., the approximation of the ligand by a bounding sphere.
\textcolor{blue}{However, the ligands are typically of a non-spherical shape which is completely ignored by these methods. 
Our approach aims to overcome this problem by taking into account the actual shape of the ligand and additionally also its evolution over time, i.e. its flexibility.}

@@

Our proposed method was tested by the biochemists who compared the results with already known trajectories of ligands, determined by their traditional approaches.
\textcolor{blue}{In the following we first mention several approaches and methods related to our solution and then describe our proposed method in detail.}
